Latitudinal gradients in biotic niche breadth vary across ecosystem types.

### 2015-09-07 manuscript/AC-reply-to-referees.tex

  \begin{refquote}

    L100-106. After reading the introduction, it is not clear why you ‘were
    also interested’ in this topic. I suggest following the advice of reviewer
    2 and moving this part and the associated results to the appendix. Please
    also then explain why this is a priori interesting and relevant to
    investigate in the context of your study.

  \end{refquote}

  \textbf{R:} After carefully considering the Associate 
  Editor and Reviewer's comments, we have opted to remove
  the by-trophic-level analyses entirely. Due to length
  restrictions, we were not able to both expand our 
  description of the motivations for and implications of
  these analyses and accommodate the other changes suggested
  by the Associate Editor and Reviewers. We chose to focus
  on ensuring that the discussion of the species-richness
  results is as clear and complete as possible, and believe
  that our manuscript is stronger and more to-the-point
  after these changes.

@@

  \begin{refquote}

    L113. To what extent do the property estimates covary with each other?
    This may be presented in an appendix.

  \end{refquote}

  \textbf{R:} As the Associate Editor likely suspected, the 
  property estimates are all strongly positively correlated. 
  As generality and vulnerability describe the two 
  components (number of prey and number of predators,
  respectively) of links per species, this is to be 
  expected. We have added a brief description of these 
  correlations to the main text (lines 122-125) as we 
  believe it will provide some context for the very similar 
  results for link density and vulnerability.

@@

  \begin{refquote}

    L116. It seems the stats to support this statement are missing.

  \end{refquote}

  \textbf{R:} We have ammended the text (lines 135-138) to support our statement.

@@

  \begin{refquote}

    It is particularly interesting in that stream systems (as opposed to
    estuarine, marine or terrestrial food webs) are more likely to be affected
    by seasonal variability – and this is the mechanism evoked as underpinning
    the latitudinal gradient. I am surprised that the authors didn’t make more
    of this point – to some degree the difference in habitats supports their
    proposed mechanism, and this could be discussed.

  \end{refquote}

  \textbf{R:} We thank the Reviewer for this suggestion, 
  and have substantially revised our discussion (lines 
  182-217) to follow it. The Reviewer will note that our 
  arguments include lakes as well as streams since both 
  systems showed effects of latitude on scaling, and both 
  experience severe seasonal variability as the Reviewer
  points out for streams. We have also added a brief note 
  to address the potential that the New Zealand stream food 
  webs mentioned by the Reviewer above might be affecting 
  our results. As described below, they are grouped in a 
  narrow band of latitude; this likely explains their 
  minimal impact on the trends we observe. We think that 
  our discussion is much stronger after these changes.

@@

    \smallskip

    Please justify why you use the proportion of species within a trophic
    level and not the absolute number of species per trophic level; at a first
    glance, this would be more intuitive for me given the focus of the study
    on scaling relationships with species richness.

  \end{refquote}

  \textbf{R:} We appreciate the Reviewer's concerns about 
  the integration of these results into the main text. Due 
  to length restrictions, we were unable to explain our 
  motivations and discuss our results more fully in the main
  text. Moreover,
  referring briefly to by-trophic-level results in the supplemental
  information repeated the problems with lack of justification and
  poor integration of the by-trophic-level results with the 
  species-richness results. To that end we have opted to remove
  these results entirely. Instead, we have prioritised enriching
  our discussion of the species-richness results, clarifying our
  methods, and including the latitudinal gradient in species-richness
  as requested below. We hope that, on balance, this restructuring has
  resulted in a clearer and more well-integrated manuscript.

@@

     \smallskip

      The start of the Discussion confused me since it presents another
      additional analysis. This analysis on the latitudinal gradient in
      species richness is interesting and important for the interpretation of
      the patterns, but should already be presented in Methods and Results
      section.

  \end{refquote}

  \textbf{R:} Given the lack of significant relationships,
  we had originally viewed this analysis as a very minor 
  part of our paper, but are happy to follow the Reviewer's 
  suggestion and integrate it in the methods and
  results along with our main analyses. We now describe 
  the methods for this analysis in lines 74-86 and 
  summarize our results in lines 125-128.

@@

    Please describe in more detail why you used the chosen habitat categories
    and provide the number of webs per habitat type also in the main
    manuscript.

  \end{refquote}

  \textbf{R:} We followed the habitat categories given by the publications which
  aggregated the food webs. Although more detailed descriptions were also provided, 
  we opted to use the set of five more general habitat types in order to have a 
  greater number of food webs in each category.

  We have added a brief explanation, and the number of webs in each ecosystem type, at lines 52-55:

@@

  \begin{refquote}

    Please re-label the y-axis on Fig. 1 to improve clarity. The top predators
    are missing in this figure although they are mentioned in the legend.

  \end{refquote}

  \textbf{R:} We thank the reviewer for pointing this out 
  and, while the inclusion of top predators specifically
  is moot since these results have been moved to the 
  supplemental information, have taken care that all labels
  are included in all figures. We have also re-labelled the
  y-axes on Fig. 2 (the old Fig. 1) to improve clarity.

### 2015-09-07 manuscript/latitudepaper.tex
Changed main text, reply

    We then repeated the above analyses to measure the 
    effect of latitude on the scaling of specialisation 
    with numbers and proportions of species at different 
    trophic levels (i.e., basal resources, intermediate 
    consumers, and top predators). As different trophic 
    levels represent different sections of the biotic niche 
    space, we expected that latitude might affect the 
    scaling of specialisation with the proportion of 
    species in each trophic level differently. The results
    of these supplemental analyses largely echo the the
    results of the species-richness analyses, and are 
    presented in~\emph{Appendix S4}.

@@

\begin{figure}[h]
% \centerline{\includegraphics*[height=.5\textheight]{Figures/by_TL/scaling_with_S/proportions/S_fitlines_nonts.eps}}
\caption{Scaling relationships for re-scaled link density, generality, and vulnerability 
relative to the species richness of a food web. Link density, generality,
and vulnerability were each re-scaled to remove the effects of latitude and ecosystem
type. As these relationships take the form of power laws, we did this by dividing the food-web
property (e.g. link density) by species richness raised to an exponent including the 
effects of latitude and, where applicable, ecosystem type and the interaction between ecosystem
type and latitude. Note that in all cases estuarine food webs were treated as the baseline 
ecosystem type, but that at most two ecosystem types had interactions between ecosystem type and
latitude retained in the best-fit model (see \emph{Results} for specifics). For each relationship, 
we show the re-scaled values (white circles) as well as the overall scaling relationship using estuarine
ecosystems as a baseline (black line, N=196 food webs). For a figure with the uncorrected values,
see Fig.~\emph{S7}, \emph{Appendix S5}.}
\label{props_v_lat}
\end{figure}

>>>>

    % We then repeated the above analyses to measure the 
    % effect of latitude on the scaling of specialisation 
    % with numbers and proportions of species at different 
    % trophic levels (i.e., basal resources, intermediate 
    % consumers, and top predators). As different trophic 
    % levels represent different sections of the biotic niche 
    % space, we expected that latitude might affect the 
    % scaling of specialisation with the proportion of 
    % species in each trophic level differently. The results
    % of these supplemental analyses largely echo the the
    % results of the species-richness analyses, and are 
    % presented in~\emph{Appendix S4}.

@@

\begin{figure}[h]
% \centerline{\includegraphics*[height=.5\textheight]{Figures/by_TL/scaling_with_S/proportions/S_fitlines_nonts.eps}}
\caption{Scaling relationships for re-scaled link density, generality, and vulnerability 
relative to the species richness of a food web. Link density, generality,
and vulnerability were each re-scaled to remove the effects of latitude and ecosystem
type. As these relationships take the form of power laws, we did this by dividing the food-web
property (e.g. link density) by species richness raised to an exponent including the 
effects of latitude and, where applicable, ecosystem type and the interaction between ecosystem
type and latitude. Note that in all cases estuarine food webs were treated as the baseline 
ecosystem type, but that at most two ecosystem types had interactions between ecosystem type and
latitude retained in the best-fit model (see \emph{Results} for specifics). For each relationship, 
we show the re-scaled values (white circles) as well as the overall scaling relationship using estuarine
ecosystems as a baseline (black line, N=196 food webs). For a figure with the uncorrected values,
see Fig.~\emph{S7}, \emph{Appendix S4}.}
\label{props_v_lat}
\end{figure}
